A flexible framework to prototype, develop and manage ubiquitous applications for intelligent environments

### 2013-10-04 trunk/ labtempo-ubicomp-projects/UbicompMiddleware/doc/PerCom2014/percom2014_conf.tex
finalization of the paper written related work removed subsection location (from SmartAndroid) corrected author names corrections

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{\IEEEauthorblockN{Matheus Erthal}
\IEEEauthorblockA{Institute of Computing\\
Federal Fluminense University\\
Rio de Janeiro, Brazil\\
Email: merthal@ic.uff.br}
\and
\IEEEauthorblockN{David Barreto}
\IEEEauthorblockA{Institute of Computing\\
Federal Fluminense University\\
Email: dbarreto@ic.uff.br}
\and
\IEEEauthorblockN{Douglas Mareli}
\IEEEauthorblockA{Institute of Computing\\
Federal Fluminense University\\
Email: dmareli@ic.uff.br}
\and
\IEEEauthorblockN{Orlando Loques}
\IEEEauthorblockA{Institute of Computing\\
Federal Fluminense University\\
Email: loques@ic.uff.br}}
% \author{\IEEEauthorblockN{Michael Shell}
% \IEEEauthorblockA{School of Electrical and\\Computer Engineering\\
% Georgia Institute of Technology\\
% Atlanta, Georgia 30332--0250\\
% Email: http://www.michaelshell.org/contact.html}
% \and
% \IEEEauthorblockN{Homer Simpson}
% \IEEEauthorblockA{Twentieth Century Fox\\
% Springfield, USA\\
% Email: homer@thesimpsons.com}
% \and
% \IEEEauthorblockN{James Kirk\\ and Montgomery Scott}
% \IEEEauthorblockA{Starfleet Academy\\
% San Francisco, California 96678-2391\\
% Telephone: (800) 555--1212\\
% Fax: (888) 555--1212}}

@@

@@

The remainder of this paper is organized as follows. First we present in Section~\ref{sec:overview} an overview of the main concepts used as a basis for the framework's development. In Section~\ref{sec:framework}, we present the framework's overall architecture and main features. We then, in Section~\ref{sec:evaluation}, present a proof of concept demonstrating the feasibility of building UbiApps using the framework and we show examples of applications that explores the key features of the IE. Finally, in Section~\ref{sec:related} we compare our approach with related work and in Section~\ref{sec:conclusion} we conclude this paper with the main remarks.
[refactor this paragraph]

@@

Most approaches focus on one aspect of UbiComp, such as security~\cite{}, location~\cite{}, context representation~\cite{}, SmartHome specific~\cite{dixon2012operating}, among others. [+++]
% The works that focus on context interpretation/context rules
% Some works focus on composing and managing context rules  %About the adaptability 
%Challenges in Access Right Assignment for Secure Home Networks - 17

[best classical references]
\cite{satyanarayanan2001pervasive}
\cite{Park2007simulation}
\cite{ranganathan2003middleware}
\cite{dey2001conceptual}
\cite{helal2005gator}

[others]
\cite{brush2011home}
\cite{edwards2001home}
\cite{garcia2010towards}
\cite{baldauf2007survey}
\cite{gu2005service}
\cite{kindberg2002system}

[only old references]

[get references from this conference]

%TODO: Gaia
Being one pioneer in the field, the Gaia middleware was designed to facilitate the construction of applications for IE.It consists of a set of core services and a framework for building distributed context aware applications. The Gaia embraces many goals that include the acquisition of context, the maintenance of hardware and software descriptions, mechanisms to find resources and to compose context rules, among others.
% FIXME: it really permits to compose context rules or just proposes?

%uses operating systems abstractions to define its architecture
%have many goals that include the provision of mechanisms of acquisition of context; composition of rules (using first logic order), 
%%maintenance of hardwares and softwares description and about applications; 
%search mechanism to find resources
%presence and execution status of resources
%interpretation mechanism that gathers and interprets information, as defined by programmers
%mvc
%The main goals of Gaia's approach are the acquisition of context by applications, the monitoring of entities location, the maintenance of hardwares and softwares descriptions, and 
%Active Spaces
%FIXME: second phrase was copied

%TODO: Context Toolkit - Dey
%also a famous approach
%we use these abstractions, the aggregation, interpretation, the widget
The Context Toolkit provides a set of abstractions for the composition of reusable components that focus on context acquisition and interpretation. [+++]

%TODO: Gator Tech
[Gator Tech]
%generic reference architecture applicable to any IE 
%middleware layers:
%physical, sensor platform (OSGi , service, knowledge, context management, application

@@

% [talk about market, which approach are already in use?] Google IS, onX

@@

@@

Context aware (or context sensitive) applications is a designation to those applications that not only are capable of knowing the context of the environment, but also react to it either by means of changing the environment or in a software level. []

[distributed systems]

Ubiquitous systems are a class of distributed systems. So, our approach incorporates well proovem distributed system concepts that are essential to disseminate and manage context information. Besides this, a distributed component model is adopted to implement the entities that live and interact in the ubiquitous environment.
% FIXME: prooven?

@@

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%              SECTION
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% FIXME: Framework Details????????????????????
\section{Framework Features}
This section describes details of the main features provided by the framework, which includes the definition of the RA (Subsection~\ref{subsec:ra}), the management services (Subsection~\ref{subsec:sgar}) and the context interpretation (Subsection~\ref{subsec:ci}). In the next section we will present the current status of the implementation.
% This section approaches the main features provided by the framework, which includes the definition of the RA (Subsection~\ref{subsec:ra}), the management services (Subsection~\ref{subsec:sgar}) and the context interpretation (Subsection~\ref{subsec:ci}). Also, based on the reference implementation, an application was built to provide a tool for the prototyping, management and visualization of applications and resources running on the IE.

@@

  [wrapper]

@@

    \subsubsection{Security Management Service}
    The security concern is always a central topic on UbiComp systems, due to the diverse aspects of security required. Some aspects are already inherited, depending on the technology adopted, e.g., the \mbox{Wi-Fi} already ensures against the registration of outside devices and the exchange of information. Thus our proposal for the the SecMS discusses requirements at a higher level, and issues at a lower level are defined according to the platform used.

@@

%   [timer can be CV]
  Another aspect of context rules is the time dimension. The definition of timers inside a rule expression can be done by subscribing the CI to a Timer RA, hence enabling the direct reference in the expression. The timer can also be relative or absolute. An absolute timer is objective: if the timer has reached some timestamp, or if it is within a window of time, then it is valid. [+++]

@@

\section{SmartAndroid}
% [android and other techs]
%We consolidated the concepts elaborated in the course of this project as the platform \mbox{\textit{SmartAndroid}}\footnote{For more information visit www.tempo.uff.br/smartandroid}. The project includes the implementation of APIs to , as well as prototyping and management interface on the Android platform.

%[David]
The conceptual framework proposed at this work was tested through a reference implementation called \mbox{SmartAndroid}~\footnote{For more information visit www.tempo.uff.br/smartandroid}. The project includes a programming API to help the developers in constructing UbiApps as well as offers basic management services to provide resource registration, discovery and localization. The SmartAndroid has been developed over Google Android platform as we believe that in a near future it will be widely used in sensors and devices embedded in furnitures and home appliances.
% [we have implemented most of the framework features, thought much more remain to be done]

@@

  \subsection{Location}

  [talk about location]

  The devices' indoor localization may be acquired through many different approaches, as an example, using radio frequency devices such as RFID tags and Wifi devices~\cite{Talukder2007Tracker} or even a smart floor~\cite{Branzel2013SmartFloor} which can locate people and objects by a pressure-sensing floor. However, the SmartAndroid doesn't implement a specific technique to locate resources, in order to keep the framework generic. The SmartAndroid API deal with location in a hi-level way, making the data acquirement transparent to it. This aspect is important to keep the system flexible to any approach, allowing the developer to use different implementations.

@@

  \begin{figure}[ht]
  \centering
  \includegraphics[width=3.4in]{figures/publish-subscribe-ar.pdf}
  \caption{Resource Agent} 
  \label{fig:ra-intercept}
  \end{figure}

  \subsection{Context Rules Composition} \label{subsec:rules}
%   [implementation of context rules]
  The SmartAndroid supports the context interpretation by providing mechanisms to create context rules. The context rules life cycle has two major phases: composition and running. The composition phase corresponds to the instantiation of a CI and of an actuator and the binding of both through subscription. The running phase is when the CI and the actuator(s) are already instantiated as the other RAs that feed the IE with context information, and they are able to exchange messages and perform their job.
  %   Support the interpretation of the context framework enables a separation of concerns, where developers abstract implementation of contextual rules and focus on the application logic.

  The composition of context rules can be done basically in two ways. The former is at the implementation level, where developers are able define their context rules in a description file or using a programming language. The latter is at end user level, where non trained people are able to define their context rules supported by the PIPA (Subsection~\ref{subsec:ipgap}).

  At the implementation level, a context rule can be described in a JSON structured file (or byte stream) following a specific format. In the first part an expression must be defined and in the second part the actuators can be defined. This expression contains conditions that are the comparison of RA's CVs with constant values (e.g., Stove.OvenTemperature = $180\,^{\circ}\mathrm{C}$), of two different CVs from the same or from different RAs (e.g., ArConditioning.Temperature $>$ OutsideSensor.Temperature), or the RA's CV alone if it is has a boolean value (e.g., Stove.IsOn). The conditions are bound to each other through the logical operators ``AND'' and ``OR'', it is also allowed the use of ``NOT'' and of parenthesis, what makes possible to compose more complex expressions by using subexpressions. The expression or subexpression may also contain a timer.

@@

  [review of capabilities]

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%              SECTION
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion} \label{sec:conclusion}
% [framework main features]
In the present paper we have proposed a framework that address some major issues that are found in the construction of ubiquitous applications, both for developers as for end users. We have focused on three challenges, that have guided the development of the framework. For dealing the the heterogeneity of devices present in the intelligent environment we have created the Resource Agent abstraction that encapsulates the mechanisms to interact with physical and also virtual resources, and includes a synchronous and asynchronous communication mechanism. The variety of context information and services is achieved by the definition of Context Interpreters that are responsible for aggregating context information, providing it in a higher level. And for the resource availability issue we have created the PIPA, which is a software that helps the prototyping and testing of ubiquitous applications through a easy to use interface, and allows the visualizations of applications running in a environment populated with real and simulated devices.

% [smartandroid]
The concepts of the framework were implemented as a middleware called SmartAndroid. The SmartAndroid provides a API for developers and a set of basic services, that enables the construction of the intelligent environment. The core management services are the resource register service, that includes a naming service; the resource discovery service, that performs searches on registered resources based on different criteria; and a resource location service, that also performs searches but based on the location and proximity of resources and/or users.

>>>>

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
% \author{\IEEEauthorblockN{Matheus Erthal}
% \IEEEauthorblockA{Institute of Computing\\
% Federal Fluminense University\\
% Rio de Janeiro, Brazil\\
% Email: merthal@ic.uff.br}
% \and
% \IEEEauthorblockN{David Barreto}
% \IEEEauthorblockA{Institute of Computing\\
% Federal Fluminense University\\
% Email: dbarreto@ic.uff.br}
% \and
% \IEEEauthorblockN{Douglas Mareli}
% \IEEEauthorblockA{Institute of Computing\\
% Federal Fluminense University\\
% Email: dmareli@ic.uff.br}
% \and
% \IEEEauthorblockN{Orlando Loques}
% \IEEEauthorblockA{Institute of Computing\\
% Federal Fluminense University\\
% Email: loques@ic.uff.br}}

% \author{\IEEEauthorblockN{Michael Shell}
% \IEEEauthorblockA{School of Electrical and\\Computer Engineering\\
% Georgia Institute of Technology\\
% Atlanta, Georgia 30332--0250\\
% Email: http://www.michaelshell.org/contact.html}
% \and
% \IEEEauthorblockN{Homer Simpson}
% \IEEEauthorblockA{Twentieth Century Fox\\
% Springfield, USA\\
% Email: homer@thesimpsons.com}
% \and
% \IEEEauthorblockN{James Kirk\\ and Montgomery Scott}
% \IEEEauthorblockA{Starfleet Academy\\
% San Francisco, California 96678-2391\\
% Telephone: (800) 555--1212\\
% Fax: (888) 555--1212}}

@@

\author{\IEEEauthorblockN{Matheus Erthal\IEEEauthorrefmark{1},
David Barreto\IEEEauthorrefmark{1},
Douglas Mareli\IEEEauthorrefmark{1} and
Orlando Loques\IEEEauthorrefmark{1}}
\IEEEauthorblockA{\IEEEauthorrefmark{1}Institute of Computing\\
Federal Fluminense University\\
Rio de Janeiro, Brazil\\ 
\{merthal, dbarreto, dmareli, loques\}@ic.uff.br}}

@@

The remainder of this paper is organized as follows. First we compare our approach with related works in Section~\ref{sec:related}, then in Section~\ref{sec:overview} we present an overview of the main concepts used as a basis for the framework's development. In Section~\ref{sec:framework}, we describe the framework's overall architecture as well as, in Section~\ref{sec:features}, its most important features. Our framework reference implementation as the assumptions, techniques and features related are discussed in Section~\ref{sec:smartandroid}. A proof of concept demonstrating the feasibility of building UbiApps using the framework is presented in Section~\ref{sec:evaluation}, showing examples of applications that explores the key features of the IE. Finally, in Section~\ref{sec:conclusion} we conclude this paper with the main remarks and future works.

%The remainder of this paper is organized as follows. First we present in Section~\ref{sec:overview} an overview of the main concepts used as a basis for the framework's development. In Section~\ref{sec:framework}, we present the framework's overall architecture and main features. We then, in Section~\ref{sec:evaluation}, present a proof of concept demonstrating the feasibility of building UbiApps using the framework and we show examples of applications that explores the key features of the IE. Finally, in Section~\ref{sec:related} we compare our approach with related work and in Section~\ref{sec:conclusion} we conclude this paper with the main remarks.
% [refactor this paragraph]

@@

% The works that focus on context interpretation/context rules
% Some works focus on composing and managing context rules  %About the adaptability 
%Challenges in Access Right Assignment for Secure Home Networks - 17

% [best classical references]
% \cite{satyanarayanan2001pervasive}
% \cite{Park2007simulation}
% 
% \cite{dey2001conceptual}
% \cite{helal2005gator}
% 
% [others]
% \cite{brush2011home}
% \cite{edwards2001home}
% \cite{garcia2010towards}
% \cite{baldauf2007survey}
% \cite{gu2005service}
% \cite{kindberg2002system}
% 
% [only old references]
% 
% [get references from this conference]

%TODO: Gator Tech
%generic reference architecture applicable to any IE 
%middleware layers:
%physical, sensor platform (OSGi , service, knowledge, context management, application

@@

@@

%%%%%%%%%%%%%%% HETEROGENEITY

%%%%%%%%%%%%%%% CONTEXT AWARENESS
%TODO: Context Toolkit - Dey
%also a famous approach
%we use these abstractions, the aggregation, interpretation, the widget
%This approach provides a starting point
% The Context Toolkit~\cite{dey2001conceptual} is a seminal work that provides a set of abstractions for the composition of reusable components that focus on context acquisition and interpretation to support rapid prototyping of context aware applications. 
The Dey and Abowd framework~\cite{dey2001conceptual}, implemented by the Context Toolkit, is a seminal work for the supporting of rapid prototyping of context aware applications and provides a set of abstractions for the composition of reusable components that focus on context acquisition and interpretation. Another pioneer in the field, the Gaia middleware~\cite{ranganathan2003middleware} was designed to facilitate the construction of applications for IE. It consists of a set of core services and a framework for building distributed context aware applications. The Gaia embraces many goals that include the acquisition of context, the maintenance of hardware and software descriptions, mechanisms to find resources and to compose context rules, among others. However, these two approaches do not have concerns on the user side and their preferences, including the prototyping and management of UbiApps at user level.
% Being one pioneer in the field, 
%uses operating systems abstractions to define its architecture
%have many goals that include the provision of mechanisms of acquisition of context; composition of rules (using first logic order), 
%%maintenance of hardwares and softwares description and about applications; 
%search mechanism to find resources
%presence and execution status of resources
%interpretation mechanism that gathers and interprets information, as defined by programmers
%mvc
%The main goals of Gaia's approach are the acquisition of context by applications, the monitoring of entities location, the maintenance of hardwares and softwares descriptions, and 
%Active Spaces
%FIXME: second phrase was copied

% DIOS++
The framework proposed in DIOS++~\cite{liu2003dios++} offers abstractions to manipulate sensors and actuators, a communication layer to manage distributed sensors and actuators, and a distributed rule engine. But concerns on temporization of context rules and creation of rules at end user level were not touched upon.

% CASS
The CASS~\cite{Park2007simulation} is a simulation rule environment that have a simple mechanism to discover possible conflicts between rules that are composed only using ``AND'' clause. However, it does not deploy the rules in a running environment.

%%%%%%%%%%%%%%%%%% PROTOTYPING
The DiaSim~\cite{bruneau2012} presents a simulator for pervasive applications, but does not support dynamic configuration of  entities in the system. Furthermore the work is not clear about the transparency between real and simulated devices. Other works on simulation of pervasive systems~\cite{barton2002ubiwise, nishikawa2006ubireal} provide a 3D visualization of the environment as in a game. It is possible to configure devices within the system and interact with them. However these works focus on interaction tests of devices and simulation of the physical environment, not worrying about context awareness.

The BASE~\cite{becker2003base} provides a minimal middleware for handling the resources by means of simple abstractions to access remote services and device-specific capabilities, but it focus only on the communication concern.
%%%%%%%%%%%%%%%%%% Specific purpose approaches
% Most approaches focus on one aspect of UbiComp, such as security~\cite{kim2010challenges}, location~\cite{}, context representation~\cite{}, SmartHome specific~\cite{dixon2012operating}, among others. [+++]
% There are works that focus on one aspect of UbiComp, such as security~\cite{kim2010challenges}, location~\cite{}

%%%%%%%%%%%%%%%%%% FOR SELLING
% [talk about market, which approach are already in use?] Google IS, onX

@@

Context aware (or context sensitive) applications is a designation to those applications that not only are capable of knowing the context of the environment, but also react to it either by means of changing the environment or in a software level. 
% TODO: citation

% [distributed systems]
Ubiquitous systems are a class of distributed systems. So, our approach incorporates well established distributed system concepts that are essential to disseminate and manage context information. Besides this, a distributed component model is adopted to implement the entities that live and interact in the ubiquitous environment.
% FIXME: prooven?

@@

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%              SECTION
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% FIXME: Framework Details????????????????????
\section{Framework Features} \label{sec:features}
This section describes details of the main features provided by the framework, which includes the definition of the RA (Subsection~\ref{subsec:ra}), the management services (Subsection~\ref{subsec:sgar}) and the context interpretation (Subsection~\ref{subsec:ci}). In the next section we will present the current status of the implementation.
% This section approaches the main features provided by the framework, which includes the definition of the RA (Subsection~\ref{subsec:ra}), the management services (Subsection~\ref{subsec:sgar}) and the context interpretation (Subsection~\ref{subsec:ci}). Also, based on the reference implementation, an application was built to provide a tool for the prototyping, management and visualization of applications and resources running on the IE.

@@

%  TODO: [wrapper]

@@

    \subsubsection{Security Management Service}
    The security concern is always a central topic on UbiComp systems, due to the diverse aspects of security required~\cite{kim2010challenges}. Some aspects are already inherited, depending on the technology adopted, e.g., the \mbox{Wi-Fi} already ensures against the registration of outside devices and the exchange of information. Thus our proposal for the the SecMS discusses requirements at a higher level, and issues at a lower level are defined according to the platform used.

@@

%   [timer can be CV]
  Another aspect of context rules is the time dimension. The definition of timers inside a rule expression can be done by subscribing the CI to a Timer RA, hence enabling the direct reference in the expression. The timer can also be relative or absolute. An absolute timer is objective: if the timer has reached some timestamp, or if it is within a window of time, then it is valid. %TODO: [+++]

@@

\section{SmartAndroid} \label{sec:smartandroid}
% [android and other techs]
%We consolidated the concepts elaborated in the course of this project as the platform \mbox{\textit{SmartAndroid}}\footnote{For more information visit www.tempo.uff.br/smartandroid}. The project includes the implementation of APIs to , as well as prototyping and management interface on the Android platform.

%[David]
The conceptual framework proposed at this work was tested through a reference implementation called \mbox{SmartAndroid}~\footnote{For more information visit www.tempo.uff.br/smartandroid}. The project includes a programming API to help the developers in constructing UbiApps as well as offers basic management services to provide resource registration, discovery and localization. The SmartAndroid has been developed over Google Android platform as we believe that in a near future it can be widely used in sensors and devices embedded in furnitures and home appliances. In addition, this option helped us to implement emulators of several devices using cheap Android cellphones an tablets.

% [we have implemented most of the framework features, thought much more remain to be done]

@@

%   \subsection{Location}
%   
% %   [talk about location]
%   
%   The devices' indoor localization may be acquired through many different approaches, as an example, using radio frequency devices such as RFID tags and Wifi devices~\cite{Talukder2007Tracker} or even a smart floor~\cite{Branzel2013SmartFloor} which can locate people and objects by a pressure-sensing floor. However, the SmartAndroid doesn't implement a specific technique to locate resources, in order to keep the framework generic. The SmartAndroid API deal with location in a hi-level way, making the data acquirement transparent to it. This aspect is important to keep the system flexible to any approach, allowing the developer to use different implementations.

  %FIXME: talk about location internally or remove this subsection

@@

  \begin{figure}[ht]
  \centering
  \includegraphics[width=3.5in]{figures/publish-subscribe-ar-en-v2.pdf}
  \caption{Resource Agent} 
  \label{fig:ra-intercept}
  \end{figure}

  \subsection{Context Rules Composition} \label{subsec:rules}
%   [implementation of context rules]
  The SmartAndroid supports the context interpretation by providing mechanisms to create context rules. The composition of context rules can be done basically in two ways. The former is at the implementation level, where developers are able define their context rules in a description file or using a programming language. The latter is at end user level, where non trained people are able to define their context rules supported by the PIPA (Subsection~\ref{subsec:ipgap}).

  The context rule life cycle has two major phases: composition and running. The composition phase corresponds to the instantiation of a CI and of an actuator and the binding of both through subscription. The running phase is when the CI and the actuator(s) are already instantiated as the other RAs that feed the IE with context information, and they are able to exchange messages and perform their job.
  %   Support the interpretation of the context framework enables a separation of concerns, where developers abstract implementation of contextual rules and focus on the application logic.

  At the implementation level, a context rule can be described in a JSON structured file (or byte stream) following a specific format, thus the CI code and the binding with the actuators can be automatically generated. The file must contain the IC definition with the rule expression, and may contain the definition of actuators too. This expression contains conditions that are the comparison of RA's CVs with constant values (e.g., Stove.OvenTemperature = $180\,^{\circ}\mathrm{C}$), of two different CVs from the same or from different RAs (e.g., ArConditioning.Temperature $>$ OutsideSensor.Temperature), or the RA's CV alone if it is has a boolean value (e.g., Stove.IsOn). The conditions are bound to each other through the logical operators ``AND'' and ``OR'', it is also allowed the use of ``NOT'' and of parenthesis, what makes possible to compose more complex expressions by using subexpressions. The expression or subexpression may also contain a timer.

@@

%   TODO: [review of capabilities]

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%              SECTION
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion} \label{sec:conclusion}
% [framework main features]
In the present paper we have proposed a framework that address some major issues that are found in the construction of ubiquitous applications, both for developers as for end users. We have focused on three challenges, that have guided the development of the framework. For dealing with the heterogeneity of devices present in the intelligent environment we have created the Resource Agent abstraction that encapsulates the mechanisms to interact with physical and also virtual resources, and includes synchronous and asynchronous communication mechanisms. The variety of context information and services is achieved by the context variable feature, which allows the creation of customized rules. These rules define Context Interpreters, that are responsible for aggregating context information, providing it in a higher level. And for the resource availability issue we have created the PIPA, which is a software that helps the prototyping and testing of ubiquitous applications through a easy to use interface, and allows the visualizations of applications running in a environment populated with real and simulated devices.

% The variety of context information and services is achieved by the definition of Context Interpreters that are responsible for aggregating context information, providing it in a higher level.

% The variety of context information and services is achieved by the context variable feature, thus allowing the creation of customized context rules that define Context Interpreters that are responsible for aggregating context information, providing it in a higher level. 
% 
% The variety of context information and services is achieved by the context variable feature. 
% 
% The context variable feature allows the creation of customized context rules, that define Context Interpreters, which allows the creation of customized rules.

% [smartandroid]
The concepts of the framework were implemented as a middleware called SmartAndroid. The SmartAndroid provides a API for developers and a set of basic services, that enables the construction of the intelligent environment. The core management services are the resource register service, that includes a naming service; the resource discovery service, that performs searches on registered resources based on different criteria; the resource location service, that also performs searches but based on the location and proximity of resources and/or users; and the security management service, that allows the customization and ensures the limits of users and applications within the environment.
